Boutiques: a flexible framework to integrate command-line applications in computing platforms

### 2018-01-19 boutiquespaper.tex
Declaration of conflicts of interests
Adjusted style for pre-print
Added Shawn as co-author, implemented reviewers' comments, drafted letter.

%%% Journal toggle; only specific options recognised.
%%% (Only "gigascience" and "general" are implemented now. Support for other journals is planned.)
\journal{gigascience}

\usepackage{graphicx}
\usepackage{siunitx}
\papercat{Paper}

@@

\title{Boutiques: a flexible framework for automated application integration in computing platforms}

@@

\author[1]{Tristan Glatard}
\author[2,3]{Gregory Kiar}
\author[2,3]{Tristan Aumentado-Armstrong}
\author[2,3]{Natacha Beck}
\author[4]{Pierre Bellec}
\author[2,3]{R\'emi Bernard}
\author[5]{Axel Bonnet}
\author[5]{Sorina Camarasu-Pop}
\author[5]{Fr\'ed\'eric Cervenansky}
\author[2,3]{Samir Das}
\author[6]{Rafael Ferreira da Silva}
\author[7]{Guillaume Flandin}
\author[5]{Pascal Girard}
\author[8]{Krzysztof J. Gorgolewski}
\author[9]{Charles R.G. Guttmann}
\author[1]{Val\'erie Hayot-Sasson}
\author[4]{Pierre-Olivier Quirion}
\author[2,3]{Pierre Rioux}
\author[10]{Marc-\'Etienne Rousseau}
\author[2,3]{Alan C. Evans}

@@

\begin{frontmatter}
\maketitle

\begin{abstract}
We present \boutiques, a system to automatically publish, integrate
and execute applications across computational platforms. \boutiques
applications are installed through software containers described in a
rich and flexible JSON language. A set of core tools facilitate the
construction, validation, import, execution, and publishing of
applications. \boutiques is currently supported by several distinct virtual
research platforms, and it has been used to describe dozens of
applications in the neuroinformatics domain. We expect \boutiques to
improve the quality of application integration in computational
platforms, to reduce redundancy of effort, to contribute to computational
reproducibility, and to foster Open Science.
\end{abstract}

\begin{keywords}
Application integration; Containers; Neuroinformatics.
\end{keywords}
\end{frontmatter}

@@

Computational platforms such as web services, portals, scientific
gateways, workflow engines and virtual research environments commonly
integrate third-party applications to enable various types of data
processing. Applications, however, are often manually and
repeatedly integrated whereas automating and
sharing this effort would improve computational
reproducibility~\cite{peng2011reproducible,Stodden1240} and contribute
to Open Science. Meanwhile, container systems such as
Docker\footnote{\url{https://www.docker.com}} and
Singularity~\cite{kurtzer2017singularity} have emerged to facilitate
the sharing and migration of software by defining immutable, reusable
execution environments.

We present \boutiques, a system to publish, integrate and execute
command-line applications across platforms (see
Figure~\ref{fig:diagram}).  In \boutiques, a command line is described
using a flexible template comprising the inputs it requires and the
outputs it produces. Inputs may be passed directly on the command line
or through configuration files. They may also be inter-dependent, for
instance mutually exclusive. Such formal descriptions, simply referred
to as \emph{descriptors}, link to a container image where the given
application is installed. \boutiques descriptors allow for automatic
application integration in platforms, and advanced validation of input
values to prevent errors.  \boutiques descriptors are intended to be
produced by scientific application developers, stored alongside their
application, indexed by common repositories, and consumed by execution
platforms.  A set of core tools facilitate the construction,
validation, import, execution, and publishing of \boutiques
descriptors.

@@

>>>>

\usepackage{graphicx}
\usepackage{siunitx}

@@

\title{Boutiques: a flexible framework to integrate command-line applications in computing platforms}

@@

\author[1]{Tristan Glatard}
\author[2,3]{Gregory Kiar}
\author[2,3]{Tristan Aumentado-Armstrong}
\author[2,3]{Natacha Beck}
\author[4]{Pierre Bellec}
\author[2,3]{R\'emi Bernard}
\author[5]{Axel Bonnet}
\author[2,3]{Shawn T Brown}
\author[5]{Sorina Camarasu-Pop}
\author[5]{Fr\'ed\'eric Cervenansky}
\author[2,3]{Samir Das}
\author[6]{Rafael Ferreira da Silva}
\author[7]{Guillaume Flandin}
\author[5]{Pascal Girard}
\author[8]{Krzysztof J. Gorgolewski}
\author[9]{Charles R.G. Guttmann}
\author[1]{Val\'erie Hayot-Sasson}
\author[4]{Pierre-Olivier Quirion}
\author[2,3]{Pierre Rioux}
\author[10]{Marc-\'Etienne Rousseau}
\author[2,3]{Alan C. Evans}

@@

\maketitle

\begin{abstract}
We present \boutiques, a system to automatically publish, integrate
and execute command-line applications applications across
computational platforms. \boutiques applications are installed through
software containers described in a rich and flexible JSON language. A
set of core tools facilitate the construction, validation, import,
execution, and publishing of applications. \boutiques is currently
supported by several distinct virtual research platforms, and it has
been used to describe dozens of applications in the neuroinformatics
domain. We expect \boutiques to improve the quality of application
integration in computational platforms, to reduce redundancy of
effort, to contribute to computational reproducibility, and to foster
Open Science.
\end{abstract}

\begin{keywords}
Application integration; Containers; Neuroinformatics.
\end{keywords}

@@

Computational platforms such as web services, portals, science
gateways, workflow engines and virtual research environments commonly
integrate third-party applications to enable various types of data
processing. Applications, however, are often manually and
repeatedly integrated in such platforms whereas automating and
sharing this effort would improve computational
reproducibility~\cite{peng2011reproducible,Stodden1240} and contribute
to Open Science. Meanwhile, container systems such as
Docker\footnote{\url{https://www.docker.com}} and
Singularity~\cite{kurtzer2017singularity} have emerged to facilitate
the sharing and migration of software by defining immutable, reusable
execution environments.

We present \boutiques, a system to publish, integrate and execute
command-line applications across platforms (see
Figure~\ref{fig:diagram}).  In \boutiques, a command line is described
using a flexible template comprising the inputs it requires and the
outputs it produces. Inputs may be passed directly on the command line
or through configuration files. They may also be inter-dependent, for
instance mutually exclusive. Such formal descriptions, simply referred
to as \emph{descriptors}, can be parsed in any programming language,
can describe command lines regardless of their implementation,
and link to a container image where the application is
installed. \boutiques descriptors allow for automatic application
integration in platforms, and advanced validation of input values to
prevent errors.  \boutiques descriptors are intended to be produced by
application developers, stored alongside their application, indexed by
common repositories, and consumed by execution platforms.  A set of
core tools facilitate the construction, validation, import, execution,
and publishing of \boutiques descriptors.

@@

\section{Competing interests}

The authors declare that they have no competing interests.
